Solar wind predictions for the Parker Solar Probe orbit

### 2017-08-16 paperMVVB/sections/content.tex
some references

\section{Introduction}

%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1962 \citep{Neugebauer1966} \hl{+ ref. to Russian measurements}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{138}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} (July 2017), having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
Until today various spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with Wind \hl{(ref.)}, ACE \hl{(ref.)} and DSCOVR \hl{(ref.)} still orbiting around the L1 point \SI{1.5}{million \km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \hl{(ref.)} or MESSENGER \hl{(ref.)}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. The Sun-nearest in situ solar wind measurements to date were obtained by the Helios mission. The in 1974 launched Helios~1 spacecraft reached distances of \SI{0.31}{\au}, Helios~2 launched two years later and approached the Sun up to \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). {\color{Red2} In comparison, the ESA Solar Orbiter mission with a planned launch in 2019 will have a closest perihelion of \SI{0.289}{\au} \hl{(ref.)}.}

@@

%definition of sw environment
%section{Solar wind}
Generally, two types of solar wind are observed in the heliosphere, slow and fast streams \citep{Neugebauer1966,Schwenn1983}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses' out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types \citep{Schwenn1983}, is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, {\color{red} such as ...intermittent reconnection at the top of streamers and pseudostreamers, coronal hole boundaries, ... (\hl{ref.}).} The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions and for quasi-stationary coronal source regions to co-rotating interaction regions \citep{Balogh1999}.
Embedded in the slow and fast solar wind streams are transient flows of coronal mass ejections (CMEs), the faster ones driving shock waves ahead \hl{(ref.)}. Their rate follows the solar activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

@@

The solar wind velocity, density and temperature depend on the state of the solar cycle \citep{Schwenn1983}. %density modulation with SSN \citep{Schwenn1983} p.~499\\
For instance the fast solar wind is correlated with the presence of polar coronal hole extensions to lower latitudes being a typical feature of the solar cycle, being the reason for the common velocity peak in the decreasing phase of the SSN, as pointed out by \citet[\hl{p.~XX, Figure~XX}]{Bothmer2007}. Therefore the solar wind velocity, density and temperature maxima exhibit time-lags to the SSN maxima.

>>>>

\section{Introduction}
%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1959 \citep{Gringauz1960,Neugebauer1966}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{138}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} (July 2017), having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
Until today various spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with WIND \hl{(ref.)}, ACE \citep{Stone1998} and DSCOVR \citep{Burt2012} still orbiting around the L1 point \SI{1.5}{million\,\km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \hl{(ref.)} or MESSENGER \hl{(ref.)}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. The Sun-nearest in situ solar wind measurements to date were obtained by the Helios mission. The in 1974 launched Helios~1 spacecraft reached distances of \SI{0.31}{\au}, Helios~2 launched two years later and approached the Sun up to \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). In comparison, the ESA Solar Orbiter mission with a planned launch in February 2019 will have its closest perihelia at \SI{0.28}{\au} \citep{Muller2013}.

@@

%definition of sw environment
%section{Solar wind}
Generally, two types of solar wind are observed in the heliosphere, slow and fast streams \citep{Neugebauer1966,Schwenn1983}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses' out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types \citep{Schwenn1983}, is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, such as intermittent reconnection at the top of helmet streamers and from coronal hole boundaries \citep{Kilpua2016}. The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions and for quasi-stationary coronal source regions to co-rotating interaction regions \citep{Balogh1999}.
Embedded in the slow and fast solar wind streams are transient flows of coronal mass ejections (CMEs), the faster ones driving shock waves ahead \citep{Gosling1974}. Their rate follows the solar activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

@@

The solar wind velocity, density and temperature depend on the state of the solar cycle \citep{Schwenn1983}. %density modulation with SSN \citep{Schwenn1983} p.~499\\
For instance the fast solar wind is correlated with the presence of polar coronal hole extensions to lower latitudes being a typical feature of the solar cycle, being the reason for the common velocity peak in the decreasing phase of the SSN, as pointed out by \citet[p.~75, Figure~3.52]{Bothmer2007}. Therefore the solar wind velocity, density and temperature maxima exhibit time-lags to the SSN maxima.
